---
title: "lab_report_knapsack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab_report_knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(42,kind="Mersenne-Twister",normal.kind="Inversion")
n<-2000
knapsack_objects<-data.frame(w=sample(1:4000,size= n,replace=TRUE),v=runif(n= n,0,10000))
library("profvis")
```

```{r setup}
library("lab6group5")
```
This vignettes contains three solutions to solve knapsack problem.Each has different computational complexity.

# 1.brute_force_knapsack
The brute force search solution is guaranteed to give a correct answer in all situations for the knapsack problems.This approach is of complexity O(2^n).
This function has two parameters:
x:a data frame which includes two columns "v" and "w".
W:an integer which means the given limit of weight.
```
brute_force_knapsack <- function(x, W) {
  #check input
  if (is.data.frame(x) == FALSE | ncol(x) != 2) {
    stop()
  }
  if (W < 0 | sum(x < 0) > 1) {
    stop()
  }
  #all combinations
  choice_matrix = matrix(0, 2 ** NROW(x) - 1, NROW(x))
  for (i in 1:2 ** nrow(x) - 1) {
    a = intToBits(i)
    for (j in 1:NROW(x)) {
      choice_matrix[i, j] = as.integer(intToBits(i)[j])
    }
  }

  #calculate weight and value
  value_matrix = matrix(0, 2 ** NROW(x) - 1, 2)
  # total weight and total value of each combination
  for (i in 1:nrow(choice_matrix)) {
    # i is row, 1:255
    total_w = 0
    total_v = 0
    for (j in 1:ncol(choice_matrix)) {
      # j is column, 1:nrow(x)
      if (choice_matrix[i, j] == 1) {
        total_w = total_w + x[j, 1]
        total_v = total_v + x[j, 2]
      }
    }
    value_matrix[i, 1] = total_w
    value_matrix[i, 2] = total_v
  }

  # get the maximum value
  max_value = 0
  tag = 0 # the row of maximum value
  for (i in 1:nrow(value_matrix)) {
    if (value_matrix[i, 1] < W & value_matrix[i, 2] > max_value) {
      max_value = round(value_matrix[i, 2])
      tag = i
    }
  }

  #return a result list
  elements = c()
  for (i in 1:NCOL(choice_matrix)) {
    if (choice_matrix[tag, i] == 1) {
      elements = append(elements, i)
    }
  }
  result = list(max_value, elements)
  names(result) <- c("value", "elements")
  return(result)

}
```
## example:
brute_force_knapsack(x= knapsack_objects[1:12,],W=2000)

### Question:
How much time does it takes to run the algorithm for n= 16 objects?
```{r}
timestart<-Sys.time()
brute_force_knapsack(x=knapsack_objects[1:16,],W=5000)
timeend<-Sys.time()
runningtime<-timeend-timestart
print(runningtime)
```
### Answer: ≈9s

# 2.knapsack_dynamic
The dynamic programming create an algorithm that can solve the knapsack problem exact by iterating over all possible values of w.
This approach is of complexity O(Wn).
This function has two parameters:
x:a data frame which includes two columns "v" and "w".
W:an integer which means the given limit of weight.
```
knapsack_dynamic <- function(x, W) {
  if (is.data.frame(x) == FALSE | ncol(x) != 2) {
    stop()
  }
  if (W < 0 | sum(x < 0) > 1) {
    stop()
  }
  x$w <- as.numeric(x$w)
  n <- nrow(x)
  m <- matrix(0, nrow = n + 1, ncol = W + 1)
  e <- matrix("", nrow = n + 1, ncol = W + 1)
  for (j in 2:(W + 1)) {
    for (i in 2:(n + 1)) {
      if (x$w[i - 1] > (j - 1)) {
        m[i, j] <- m[i - 1, j]
        e[i, j] <- e[i - 1, j]
      } else{
        if (m[i - 1, j] >= m[i - 1, j - x$w[i - 1]] + x$v[i - 1]) {
          m[i, j] <- m[i - 1, j]
          e[i, j] <- e[i - 1, j]
        } else{
          m[i, j] <- m[i - 1, j - x$w[i - 1]] + x$v[i - 1]
          e[i, j] <-
            paste(e[i - 1, j - x$w[i - 1]], i - 1, sep = " ")
        }
      }
      i <- i + 1
    }
    j <- j + 1
  }
  e <-
    as.numeric(unlist(strsplit(e[which(m == max(m))][1], split = " ")))
  l <-
    list(value = round(max(m)), elements = e[which(is.na(e) == FALSE)])
  return(l)
}

```
## example:
knapsack_dynamic(x= knapsack_objects[1:12,],W=2000)

### Question: How much time does it takes to run the algorithm for n= 500 objects?
```{r}
timestart<-Sys.time()
knapsack_dynamic(x=knapsack_objects[1:500,],W=5000)
timeend<-Sys.time()
runningtime<-timeend-timestart
print(runningtime)
```
### Answer: ≈8s

# 3.greedy_knapsack
The greedy heuristic is to use the a heuristic or approximation for the problem. This algorithm will not give an exact result (but it can be shown that it will return at least 50% of the true maximum value). It will reduce the computational complexity considerably to O(nlogn).
This function has two parameters:
x:a data frame which includes two columns "v" and "w".
W:an integer which means the given limit of weight.
```
greedy_knapsack <- function(x, W) {
  #check input
  if (is.data.frame(x) == FALSE | ncol(x) != 2) {
    stop()
  }
  if (W < 0 | sum(x < 0) > 1) {
    stop()
  }
  #initialization
  unit_value = matrix(0, NROW(x), 4)
  #unit_value
  for (i in 1:NROW(x)) {
    unit_value[i, 1] = x[i, 2] / x[i, 1]
    unit_value[i, 2] = x[i, 2] # value
    unit_value[i, 3] = x[i, 1]
    unit_value[i, 4] = i
  }
  unit_value = unit_value[order(unit_value[, 1]),]
  #result
  elements = c()# items that has been selected
  max_value = 0
  total_weight = 0
  for (i in NROW(unit_value):1) {
    max_value = max_value + unit_value[i, 2]
    total_weight = total_weight + unit_value[i, 3]
    elements = append(elements, unit_value[i, 4])
    if (total_weight + unit_value[i - 1, 3] > W) {
      break
    }
  }
  result = list(round(max_value), elements)
  names(result) <- c("value", "elements")
  return(result)
}
```
## example:
greedy_knapsack(x= knapsack_objects[1:12,],W=2000)

### Question: How much time does it takes to run the algorithm for n= n= 1000000 objects?
```{r}
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 1000000
knapsack_objects <-
     data.frame(w = sample(1:4000, size = n, replace = TRUE),
                v = runif(n = n, 0, 10000))
timestart<-Sys.time()
greedy_knapsack(x=knapsack_objects[1:1000000,],W=5000)
timeend<-Sys.time()
runningtime<-timeend-timestart
print(runningtime)
```
### Answer: ≈36s

# This is a comparison of the speed of three functions.
```{r}
profvis({
brute_force_knapsack(x= knapsack_objects[1:12,],W=2000)
knapsack_dynamic(x= knapsack_objects[1:12,],W=2000)
greedy_knapsack(x= knapsack_objects[1:12,],W=2000)
})
```

### Question: What performance gain could you get by trying to improving your code?

### Answer: Brute force search has the best accuracy but also has the most complicated complexity. The greedy heuristic can solve largest data with easier complexity, but can not give the accurate solution. The dynamic programming has the balance of accuracy and computational complexity.
